---
title: "An Introducion to QCD"
author: 
  - Sanghee Kim
  - Sumanta Basu
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    toc: yes
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{An Introduction to QCD}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

`QCD` is a package that solves penalized quantile regression via exact coordinate descent method. The penalties considered are LASSO, SCAD, and MCP. Note that QCD algorithms for SCAD and MCP are experimental. 

This vignette describes basic usage of functions related to $\ell_1$ penalized quantile regression model in `QCD` package in R.

`QCD` mainly solves the following problem. Given data points $(x_1, y_1), \ldots, (x_n, y_n)$, where $y_i \in \mathbb{R}$ is a numerical response variable, and $x_i \in \mathbb{R}^p$ is a $p$-dimensional covariate,

$$\arg \min_{\beta \in \mathbb{R}^p} \sum_{i=1}^n \rho_\tau \left(y_i - x_i^\top \beta \right) + \lambda \sum_{j=1}^p \left| \beta_j \right|$$
where $\rho_\tau \left(u\right):= u \left(\tau-\mathbf{I}(u<0)\right)$ is the check loss function, and $\lambda$ is a penalty parameter to be chosen in a data-driven fashion. $\lambda$ could be a grid of values covering the entire range of possible solutions.


## Installation

You can download the `QCD` package from Github.

```{r, message=FALSE}
library(devtools)
devtools::install_github("sangheekim96/QCD")
```

## Quick Start: QCD

The purpose of this section is to give users a general sense of the package regarding QCD. We will briefly go over the main functions, basic operations and outputs. `QCD` can be loaded using the `library` command:

```{r}
library(QCD)
```

We create a data set for illustration:

```{r}
set.seed(1)
data <- generate.data(n = 30, p = 30, tau = 0.5, signal = 1)
x <- data$X
y <- data$Y
```

This function generates simulation data based on Peng and Wang (2015). Note that in our work, we do not consider intercept and we can change the value of true beta through signal argument. 

With single value of $\lambda$, we can fit the $\ell_1$ penalized quantile regression using 'qcd.lasso.fit'.

```{r}
qr.lasso <- qcd.lasso.fit(x = x, y = y, 
                          tau = 0.5, lambda = 0.01,
                          thresh = 1e-06, maxit = 100)
qr.lasso
```

We can fit the SCAD and MCP penalized quantile regression using 'qcd.scad.fit' and 'qcd.mcp.fit'. These functions are experimental.

```{r}
qr.scad <- qcd.scad.fit(x = x, y = y, 
                        tau = 0.5, lambda = 0.01,
                        a = 2.2,
                        thresh = 1e-06, maxit = 100)
qr.scad
```

```{r}
qr.mcp <- qcd.mcp.fit(x = x, y = y, 
                      tau = 0.5, lambda = 0.01,
                      a = 2.2,
                      thresh = 1e-06, maxit = 100)
qr.mcp
```

Users are encouraged to utilize 'qcd.path' function to solve penalized quantile regression problems. A pathwise scheme 'warm start' is used as a default.

```{r}
## Create lambda grid
upper <- 2; lower <- -6
lambda <-  2^seq(upper, lower, by = -0.2)

## warm start version
qr.lasso.warm = qcd.path(x = x, y = y, tau = 0.5,
                         funname = "LASSO", lambda = lambda,
                         nudge = FALSE, 
                         thresh = 1e-06, maxit = 10000)

qr.lasso.warm

## warm start and nudge version
set.seed(1)
qr.lasso.warm.nudge = qcd.path(x = x, y = y, tau = 0.5,
                               funname = "LASSO", lambda = lambda,
                               nudge = TRUE, nudgesd = 0.2, 
                               thresh = 1e-06, maxit = 10000)

qr.lasso.warm.nudge
```


